### 2016-09-09 response.md
Update response.md

We thank the referee for the very thorough analysis and for the detailed suggestions that helped to improve the analysis and the paper overall. In the following, we provide the original comment by the referee, followed by our response.

The major changes in the manuscript have been marked in red boldface. We note that the reorganization of Section 4 (Major Issue I) is not marked. There has also been a mild amount of reordering of tables and figures as a result. We have added a single figure (Figure 11) that shows the fraction of correctable galaxies distributed over the sample, and a short amount of explanatory text in Sections 4.2 and 8.

==========
Major Issue I: Reorganization of Section 4
------------------------------------------

***I spent an inordinate amount of time trying to piece together the analysis of Section 4 before discovering that important information was given later (in Section 4.2) rather than earlier (in Section 4.1).***

***I believe that Section 4.2 should come first: present the FERENGI sample, and discuss the scheme used to select those data that are to be analyzed to determine the debiasing correction. Then the details of the statistical model for debiasing (currently in Section 4.1) should come next.***

***Section 4.3 is fine where it is placed. (However, see minor comment 10 below, which actually is not quite as minor as the other comments.)***

Section 4 has been re-organized in the manner suggested by the reviewer. The authors agree this structure will be more clear to the readers.

------------------------------------------
Major Issue II: Statistical Model
------------------------------------------

***Due to the length of Major Issue II, it is placed at the end of this response.***

We used the referee's input to review and revise the fitting of the artificially-redshifted images (from FERENGI) and construction of the debiasing equation for the Hubble data, as follows:

We tested a variety of models for fitting f/f_0, including linear, quadratic, cubic, and exponential dependences on redshift; for each of these functional forms, we also tested combinations using either a constant zeta parameter or a zeta function that linearly depends on surface brightness. We compared the relative quality of fits using the Akaike information criteria (AIC); for each case, we found no significant differences in the results based on the choice of model. Full details and code for these tests can be viewed in this Python notebook: https://github.com/willettk/gzhubble/blob/master/python/notebooks/zeta_models_compared.ipynb.

Of all the models tested, we believe the best model is an exponential fit for f_features vs redshift using a linear dependence on zeta. This decision is based on the following points:

1) While the FERENGI data can be fit relatively well using any of the above models, the exponential is ideal because it is the only form whose limiting behaviour can be set to approach 0 and 1. This is a key requirement for the debiased votes so that they can be interpreted as normalized fractions. Furthermore, the results of the AIC test show that the exponential fit is either better than, or had similar support to the other models tested.

2) We choose a linear equation to represent zeta as a function of surface brightness. In terms of goodness-of-fit, the AIC test showed both models were of equal quality. While this and the small slope in zeta shows that there is no significant surface brightness dependence in the GZH data, we still opt to keep the more general, linear form. As similar debiasing methods are planned for future GZ projects with a greater quantity of calibration images (and for which there may exist stronger dependencies on surface brightness), keeping the general form for this project will allow for a closer comparison of the two iterations of the method. We have added an emphasis on this reasoning in the manuscript so as not to lead the readers to the conclusion that we are claiming a strong surface brightness depedence detection in this run.

Finally, we have modified Equation 2 to match the normalization of Equation 4 (as suggested in minor comment 4).

-----------------------------------
Minor Comment 1
-----------------------------------

***1) Section 3.2: what checking did the authors do to indicate that giving all anonymous classifications a weight of 1 is valid?***

From a strict sense, it is impossible to tell a priori if giving the anonymous classifiers a weight of 1 is valid (since the grouping within the anonymous data is unknown). However, there are large amounts of data from similar projects that support this approach. For the Galaxy Zoo: CANDELS project (Simmons et al. 2016, submitted), which is nearly identical in design to Galaxy Zoo: Hubble and shares a large fraction of the same user base, the distribution of the classifier consistency for both the logged-in and non-logged-in populations peak at the same value (kappa = 0.7). The anonymous users have slightly broader wings, with slightly more IP addresses at the lowest and highest levels of consistency, but the overall behavior is similar.

The roughly 50-50 split between logged-in users and non-logged-in IP addresses is typical of the ratio for Zooniverse crowdsourced projects as a group. Similar results have been published for the Planet Hunters project (Schwamb et al. 2012) and have been seen in unpublished data for Pulsar Hunters (https://www.zooniverse.org/projects/zooniverse/pulsar-hunters).

ATTACH: Figure from https://cloud.githubusercontent.com/assets/2357126/17576650/cfc59fbe-5f29-11e6-99b7-c39a43c013b6.png

-----------------------------------
Minor Comment 2
-----------------------------------

***2) Figure 6: the authors should comment (unless I missed the comment) on why the bivariate distribution of the FERENGI galaxies in (mu,z) does not match the bivariate distribution for the main sample.***

The FERENGI sample is biased in including examples of galaxies that have lower surface brightnesses than any images found in the real Hubble data, simply because the local Universe has no examples of galaxies at the high surface brightness end. This is a combination of both a finite-volume sample at low redshift in addition to some evolution in the stellar populations and size-magnitude relation over cosmic time.  We deal with this issue partially by applying the evolutionary correction described in Section 2.3.2, but this still results in a sample that spans (but does not reproduce) the surface brightness-redshift distribution of the real data. The mismatch does not affect the overall calibration accuracy as we are not correcting the population as a whole, but only galaxies at particular bins of surface brightness and redshift.

We have added a paragraph in Section 4 for clarification.

-----------------------------------
Minor Comment 3
-----------------------------------

***3) In parts of Section 4.1, there is a confusing mix of f_mu,z and f_features, which appear to represent the same thing. Some clean-up (or additional explanation as to why they are different quantities) would be useful here.***

The main difference between Equations 2 and 4 is that the former is applied to the FERENGI calibration images and used to fit the normalization and zeta parameters. The latter uses the parameters to apply corrections to individual galaxies from the true data, and is the form implemented throughout the text.

We have added text to explicitly mark the difference between the two equations in what is now Section 4.2.

-----------------------------------
Minor Comment 4
-----------------------------------

***4) I probably should have mentioned this above, but why are the authors not working with equation (4) as the basis of their regression model?***

Equation 2 was originally left unbounded because the raw data being fit was inherently bounded between vote fractions of 0 and 1. However, the original version of Equation 4 could produce vote fractions greater than 1 for certain combinations of z and f_0 if left in the same form as Equation 2. We have realized that the difference in normalization, while having only a small effect on the fitting of the data, is inconsistent and may be confusing to a reader. Therefore, we have changed Equation 2 to match Equation 4 (see response to Major Comment II, as well as Minor Comment 3). All data in the tables and catalog have been updated to account for these small changes.

----------------------------------
Minor Comment 5
----------------------------------

***5) Figure 8: "-10 < log(xi) < 10" should be "-10 < xi < 10"***

Correct, thank you. This typo has been edited in the caption (now Figure 10).

---------------------------------
Minor Comment 6
----------------------------------

***6) How many data points are there in Figure 8? Obviously more than 37, but fewer than 37 x 8.***

The final version of this figure (now Figure 10; changed slightly based on the new form of Equation 4) has 28 unique galaxies, each with multiple images that have been artificially redshifted in between three to eight bins.

----------------------------------
Minor Comment 7
----------------------------------

***7) Section 4.2: "The boundaries [of the convex hull] are then adjusted until the contamination from both groups is minimized." If I look at Figure 9, it would appear that the correctable and lower-limit regions can be perfectly separated (using a tree, for instance). Am I wrong about this?***

The regions in Figure 9 (now Figure 7) can be nearly perfectly separated in f_features-redshift space, but there are a few regions near the edges in which there is overlap between the two populations. This is affected slightly in the figure by the choice of bin size and location. However, our analysis makes the assumption that a true/physical separation should be a smooth function of z-mu-f. Implementing a convex hull in this space created a very small volume near the edge of the inner convex hull that had overlapping positive and negative data points; we optimize the overall accuracy of the hull by adjusting the outer boundaries as described in what is now Section 4.1.

Use of a different method such as a tree could potentially prove to be a superior separator. However, the new normalization technique used in this draft results in very clean separation based on the convex hull, with only a single adjustment (see https://github.com/willettk/gzhubble/blob/master/python/creating_debiased_catalog/STEP_3_ferengi_smooth_function.ipynb). Further optimization would have an extremely minimal effect on the overall results.

----------------------------------
Minor Comment 8
----------------------------------

***8) Section 4.2: "In each unclassifiable (shaded blue) z,mu bin, the spread of intrinsic values for f_features,z=0.3 for four ranges of observed f_features is computed." Where did the "four" come from? More explanation needed.***

We chose to evaluate the distribution of f_features in 5 (not 4; a typo in the earlier manuscript has been corrected) bins to match some of the common ranges of vote fractions used in previous GZ morphological studies. The bin size is chosen to be roughly twice the size of the typical scatter for any individual galaxy (which has been measured using repeat image classifications in several phases of Galaxy Zoo), and so any finer grained intervals will begin to simply fit the noise.

Some clarification has been added to the text at the end of Section 4.1.

----------------------------------
Minor Comment 9
----------------------------------

***9) Why 80% of the data (orange bars, Figure 10)? This is an ad hoc choice that doesn't correspond to any typical statistical practice (except perhaps a 10% trimmed mean, but the authors are not computing means here). And how is this information applied in Figure 11, middle panel? This is not clear.***

The authors agree with the referee that a more typical statistical range would be appropriate. We have changed the range to upper and lower 1 sigma limits, or the inner 68% of the data. The figure (now Figure 8) has been modified and text added to the end of Section 4.1.

The edges of the 1-sigma range in each bin are used to determine the lower and upper limits for the debiased vote fractions of f_features. In the middle panel of what is now Figure 9, we show only the lower limits.

----------------------------------
Minor Comment 10
----------------------------------

***10) Section 4.3: "The goodness-of-fit was evaluated in each bin, for each Task, using a normalized chi^2 metric." The authors need to provide the formula for this; the numbers don't make sense, if one assumes "normalized" is the same
as "reduced chi^2" (where the typical value for a good fit is approximately 1).***

We agree that the normalized chi-squared is not the ideal metric for evaluating the goodness-of-fit here. Rather than comparing the fits of the polynomials for Task 1 vs the rest of the tasks, we compare the available data per bin. For Task 1, an average of 59 points were fit in each z-mu bin (now Figure 7). Fitting parameters for any other task is significantly limited by the number of galaxies in each z-mu bin; the higher-order tasks have an average of only 9 galaxies per bin. This is driven by the fact that the tasks do not meet the minimum threshold of 10 individual classifications (this threshold is exceeded for every galaxy in Task 1, however). This simpler evaluation should make it more evident to the reader that the FERENGI data was not well-fit for questions beyond Task 1.

As this is not a key product of the paper, we have removed Figures A5 and A6 from the manuscript to avoid confusion and extraneous discussion. Additional text has been added at the end of Section 4.3.

----------------------------------
Minor Comment 11
----------------------------------

***11) Section 7.3: "external pulblications" -> "external publications"***

Thank you, this typo has been fixed in the text.

-------------------------------
***Major Issue II***

***The authors currently apply a two-step sequential linear regression model. The first step is to fit***

***f = f_o exp[-(z-z_o)/xi]***

***which is equivalent to the fitting a linear regression after alog-transformation:***

***log(f) = log(f_o) + z_o/xi - z/xi = beta_0 + beta_1 z .***

***(Important note: there was a minus sign left out of the exponential terms in equations (2) and (4), no? The plotted results of Figure 7 don't make sense otherwise.) The derived xi's for each galaxy are then placed into another regression as the response, with surface brightness mu as the predictor, and the linear model***

***log_10(hat_xi) = xi_0 + xi_1 mu***

***is fit. For the second linear model, the xi_1 term is not actually statistically significant, but the authors sweep this under the rug and simply use the point estimate as is to drive their debiasing correction. (I did combine the steps into one linear regression model with interaction but discovered that the model could not be fit directly. It could perhaps be fit via iteration but I did not explore this further.)***

***The model appears to be a bit contrived, and there is no evidence presented that the model actually generates good results [other than a vague mention of chi^2 = 0.04 in Section 4.3 that may or may not be applicable here]). Since there is no physical motivation for the model, I would suggest two alternatives that are simpler and may perform just as well or better. (At a minimum I would want to see a metric of fit defined [*] and computed for the current model and the alternatives, if for no other reason to show that the current model is just as good or better than the others. If it is "just as good," I'd go with a simpler alternative.) It may seem that I am asking for too much here, but since the whole goal of this paper is to provide a catalog with good estimates of f_features that one can use to select galaxy samples, the authors need to do more to show that their debiased values of f_features are actually valid! (Estimates of uncertainty would help too; the models below would be able to provide them.)***

***[*] Is the normalized chi^2 considered a metric of fit? It is not completely clear. See minor comment 10 below. Perhaps the mean-squared error would work better here; it is a standard estimator.***

***1) Linear regression, with interactions; perhaps***

***log(f/f_o) = beta_0 + beta_1 z + beta_2 mu + beta_12 z mu ,***

***although one should check other tranformations (e.g., sqrt(f/f_o), log(mu), etc.)***

***2) Forego parametrization, and apply a regression tree (or its related cousin random forest regression) to log(f/f_o)(mu,z). Play with the tree settings to achieve a proper trade-off between bias and variance (i.e., to make sure the nodes are numerous enough to capture the true dependence of log(f/f_o) on mu and z, but not so numerous that each node only has a few galaxies, making the estimates of log(f/f_o) unduly noisy). A good book to use to learn about regression trees (and linear regression too) is "An Introduction to Statistical Learning" by Gareth James et al. The authors should not worry about the "Applications in R" part, although they might find doing the analysis in R (assuming it is not already done there) might be simplest. Note that a PDF of the book is available for free from James' website.***

***Another issue in Section 4.1 is the statement "The combination of all such parameters forms a high-dimensional space, and it is not clear how to separate this into individual effects." (1) I'm not sure why the individual effects are important here: the goal is to derive a debiasing correction, not to understand how the debiasing correction works. (2) There are many methods that one can use to deal with higher-dimensional regression problems, such as PCA regression, the Lasso, best subset selection, etc.; the book mentioned above gives more details on these. I'm not going to demand that the authors work with the high-dimensional data now, but they should begin to learn these statistical methods for future work.***